Switch equivalence and weak prime domains for fusions

### 2023-12-12 fusione.tex
pulizia

\newcommand{\pgfBox}{
  \begin{pgfonlayer}{background} 
    \fill[blue!2,thick,draw=black!50,rounded corners,inner sep=3mm] ([xshift=-1.5pt,yshift=-1.5pt]current bounding box.south west) rectangle ([xshift=1.5pt,yshift=1.5pt]current bounding box.north east);
  \end{pgfonlayer}
}
\usepackage{scalerel}
\newcommand{\smallmin}{\scaleobj{0.6}{-}}
\newcommand{\Deltamin}{\Delta^{\hspace{-2pt}\begin{matrix}
			\smallmin\\
		\end{matrix}} \hspace{-1pt}}
\newcommand{\Rrel}[1]   {\stackrel{{#1}}{\Longrightarrow}}
\newcommand{\oa}{\overline a}
\newcommand{\ob}{\overline b}
\newcommand{\oc}{\overline c}
\newcommand{\od}{\overline d}
\newcommand{\rec}{\emph{rec}}
\newcommand{\fn}[1]{{\mathtt{fn}}(#1)}
%\usepackage{latexsym}
\usepackage{stmaryrd}
\def\encodep#1{\llfloor#1\rrfloor}

@@

\begin{definition}
	Let $(\X, \R)$ be a DPO-rewrityng system, with $\X$ $\mathcal{M}$-adhesive. The \emph{DPO-derivation graph} of $(\X, \R)$ is the (large)  directed graph $\gpo$ having as vertices the objects of $\X$ and in which an edge between $G$ and $H$ is a direct derivation $\dder{D}\colon G\Mapsto H$.	A \emph{derivation} $\der{D}$ between two objects $G$ and $H$ is a path between them in $\gpo$.
\end{definition}

@@

@@

\begin{definition}Let $(\X, \R)$ be a DPO-rewriting system,  an \emph{abstraction equivalence} between two derivations $\der{D}$ and $\der{D'}$ with the same length and such that $r(\der{D})=r(\der{D}')$, is a family of isomorphisms $\{\phi_X\}_{X\in \Deltamin(\der{D})}$ such that, for every $i\in [0, \lgh(\der{D})]$ the following diagram commutes
	\[\xymatrix@C=40pt{G'_i&D'_i \ar[r]^{g'_i} \ar[l]_{f'}&G'_{i+1}\\  L_i \ar[u]^{n'_i} \ar[d]_{n_i}& K_i \ar[u]^{k'_i} \ar[d]_{k_i} \ar[r]^{r_i} \ar[l]_{l_i} &R\ar[u]^{h'_i} \ar[d]_{h_i}\\G_i \ar@/_.45cm/[uu]_(.35){\phi_{G_i}}|\hole&D_i\ar[l]^{f_i}\ar@/_.45cm/[uu]_(.35){\phi_{D_i}}|\hole \ar[r]_{g}&G_{i+1}\ar@/_.45cm/[uu]_{\phi_{H_i}}}\]

@@

Given DPO-rewriting system $(\X, \R)$,  we have already noted in \Cref{rem:func} that a derivation $\der{D}$  determines a diagram $\Delta(\der{D})$ in $\X$. We can then wonder if such a diagram has a colimit. Clealry if $\der{D}$ is the empty derivation $G$ then a colimit for $\Delta(\der{D})$ is simply the object $G$. More generally, we have the following result.

\begin{lemma}\label{lem:colim}
	Let $\X$ be an $\mathcal{M}$-adhesive category and $(\X, \R)$ a left-linear DPO-rewriting system over it. The following hold true.
	\begin{enumerate}
		\item For every derivation $\der{D}=$ of length $n$ and for every $j\in[0,n-1]$, the outer boundary of following diagram is a pushout,
		\[\xymatrix@C=30pt{D_{j} \ar[r]^{g_j}\ar@{>->}[d]_{f_j}& G_{j+1} \ar[r]^-{\hat{\iota}_{G_{j+1}}} \ar@{>->}@/^.3cm/[ddl]^{\iota'_{G_{j+1}}} & \lpro \der{\hat{D}}\rpro \ar[dd]^{p_2} \\ G_j \ar[d]_{\iota'_{G_j}}\\ \lpro \der{D}' \rpro \ar[rr]_{p_1}  &&\tpro{D} }\] 
		where $\der{D}':=\{\dder{D}_i\}_{i=0}^{j-1}$, $\der{\hat{D}}:=\{\dder{D}_i\}_{i=j+1}^n$ and the two arrows $p_1\colon \lpro\der{D}' \rpro\to \tpro{D}$, $p_2\colon \lpro \der{\hat{D}}\rpro \to \tpro{D}$ are induced by the cocones $(\tpro{D}, \{\iota_{X}\}_{X\in \Delta(\der{D}')})$ and $(\tpro{D}, \{\iota_{X}\}_{X\in \Delta(\der{D}')})$, respectively.
		\iffalse 
		\item if $\der{D}$ is a derivation of length one, whose unique element is given by the following diagram
	\[\xymatrix{L \ar[d]_{m}& K \ar[d]^{k}\ar@{>->}[l]_{l} \ar[r]^{r} & R \ar[d]^{h} \\G & \ar@{>->}[l]^{f} D \ar[r]_{g}& H }\]	
		Then the colimiting cocone $(\tpro{D}, \{\iota_X\}_{X\in \Delta(\der{D})})$ exists, moreover the following square is an $\mathcal{M}$-pushout;
		\[\xymatrix{ D\ar[r]^{g} \ar@{>->}[d]_{f} & H \ar[d]^{\iota_H} \\G \ar[r]_{\iota_G} &  \tpro{D}}\]
		\fi

		\item  If $\der{D}$ is a derivation from $G$ to $H$, then the diagram $\Delta(\der{D})$ has a colimit $(\tpro{D}, \{\iota_X\}_{X\in \Delta(\der{D})})$ such that $\iota_H$ belongs to $\mathcal{M}$.

		\item Let $\der{D}$ be the concatenation $\der{D}_1\cdot \der{D}_2$ of two derivations $\der{D}_1=\{\dder{D}_{1,i}\}_{i=0}^{n_1}$ between $G$ and $H$ and $\der{D}_2=\{\dder{D}_{2,j}\}_{j=0}^{n_2}$ between $H$ and $T$,  then the colimiting cocone $(\tpro{D}, \{\iota_X\}_{X\in \Delta(\der{D})})$ exists too and there is a pushout square
		\[\xymatrix{H\ar[r]^-{\iota_{2, H}} \ar@{>->}[d]_-{\iota_{1, H}} & \tproi{D}{2} \ar[d]^{p_2}\\  \tproi{D}{1} \ar[r]_{p_1}& \tpro{D}}\]
		where $(\tproi{D}{1}, \{\iota_{1, X}\}_{X\in \Delta(\der{D}_1)})$ and $(\tproi{D}{2}, \{\iota_{2, X}\}_{X\in \Delta(\der{D}_2)})$ are the colimiting cocone for $\Delta(\der{D}_1)$ and $\Delta(\der{D}_2)$, respectively.
	\end{enumerate}
\end{lemma}
\begin{remark}\label{rem:cof}
	contenuto...
\end{remark}

\begin{remark}\label{rem:zero}
There is a version of point $4$ of \Cref{lem:colim} for $j=0$ which follows directly from points $1$ and $3$. Indeed, $\der{D}=\der{D}_1 \cdot \der{D}_2$ where $\der{D}_1=\{\dder{D}_0\}$ and $\der{D}_2=\{\dder{D}_i\}_{i=1}^n$. By point $1$ and $3$ of \Cref{lem:colim}, the two halves of the rectangle below are pushouts. By \Cref{lem:po1} the whole diagram is then a pushout too.
	\[\xymatrix@C=30pt{D_{0} \ar[r]^{g_0}\ar[d]_{f_0}& G_{1} \ar[d]^{\iota_{1, G_1}} \ar[r]^-{\iota_{2,G_{1}}} & \lpro \der{D}_2\rpro \ar[d]^{p_2} \\ G \ar[r]_{\iota_{1,G}}& \lpro \der{D}_1 \rpro \ar[r]_{p_1}  &\tpro{D} }\]
Notice that this entails that the square below is also a pushout. 
\[\xymatrix@C=25pt{D_0\ar[r]^-{\iota_{2, D_0}}  \ar[d]_{\iota_{1, D_0}}&  \lpro \der{D}_1 \rpro\ar[d]^{p_2}\\ \lpro \der{D}_1 \rpro \ar[r]_{p_1}  &\tpro{D} }\]
\end{remark}

\begin{proof}\begin{enumerate}
		\item We can start noticing that $f$, being the pushout of $l$ is in $\mathcal{M}$ and thus it admits a pushout along $g$, giving us the diagram 
		\[\xymatrix{L \ar[d]_{m}& K \ar[d]^{k}\ar[l]_{l} \ar[r]^{r} & R \ar[d]^{h} \\G \ar@/_.2cm/[dr]_{q}& \ar[l]^{f} D \ar[r]_{g}& H  \ar@/^.2cm/[dl]^{p}\\ & P}\]
		Clearly, precomposing $p$ and $q$ with the arrows appearing in $\dder{D}$, we can extend the pushout $(P, p, q)$, to a cocone $(P, \{j_X\}_{X\in \Delta(\der{D})})$ on $\Delta(\der{D})$. Let $(C, \{c_X\}_{X\in \Delta(\der{D})})$ be another cocone, for $\Delta(\der{D})$, in particular we have
		\begin{align*}
			c_H\circ g & = c_D \\&= c_G\circ f
		\end{align*}
		Hence, there exists a unique $c\colon P\to C$ such that $c\circ q= c_G$ and $c\circ p =c_H$. But these identities are enough to deduce that $(P, \{j_X\}_{X\in \Delta(\der{D})})$  is colimiting.
		\item Let us proceed by induction.
		\begin{itemize}
			\item  If $n=0$, then the claim follows immediately from point $1$, noticing that $\iota_H$ is the pushout of $f$ and thus an element of $\mathcal{M}$.
			\item Let $\der{D}$ be $\{\dder{D}_i\}_{i=0}^n$ and suppose that $n\geq 1$. Let also $\der{D}'$ be $\{\dder{D}_i\}^{n-1}_{i=0}$ and $\rho_n=(l_n, r_n)$ be the rule applied in $\dder{D}_n$. The pushout of $l_n$ is the arrow $f_n\colon D_n\to G_n$ is in $\mathcal{M}$ and, by inductive hypothesis, $\iota_{G_{n}}\colon G_{n}\to \lpro \der{D}'\rpro$ is in $\mathcal{M}$ too. Thus, we can consider the diagram below, having a pushout as its lower half.
			\[\xymatrix{L_n \ar[d]_{m_{n}}& K_{n} \ar[d]^{k_{n}}\ar[l]_{l_{n}} \ar[r]^{r_{n}} & R_{n} \ar[d]^{h_n} \\G_{n} \ar[d]_{\iota'_{G_{n}}}& \ar[l]^{f_n} D_n \ar[r]_{g_n}& H  \ar[d]^{q}\\ \lpro \der{D}' \ar[rr]_{p}\rpro && P}\] 
			Notice that, as in the point above, the arrow $q\colon H\to P$ is the pushout of an element in $\mathcal{M}$, therefore it is enough to show that the diagram so constructed provides a colimiting cocone for $\Delta(\der{D})$.

			Let $(C, \{c_X\}_{X\in \Delta(\der{D})})$ be a cocone, since $\Delta(\der{D}')$ is a subdiagram of $\Delta(\der{D})$, we get another cocone $(c, \{c_X\}_{X\in \Delta(\der{D}')})$ which induces an arrow $c'\colon \lpro \der{D}' \rpro \to C$ such that
			\begin{align*}
				c'\circ \iota_{G_n} \circ f_n &=c_{G_n} \circ f_n\\&= c_{D_n}\\&= c_{H}\circ g_n
			\end{align*}
			Therefore the arrows $c'$ and $c_H$ induce a morphism $c\colon P\to C$ and the thesis now follows at once.
		\end{itemize}
		\item  As a first step, notice that $(\tpro{D}, \{\iota_X\}_{X\in \Delta(\der{D}_1)})$ and $(\tpro{D}, \{\iota_X\}_{X\in \Delta(\der{D}_2)})$ are cocone on, respectively, $\Delta(\der{D}_1)$ and $\Delta(\der{D}_2)$. Hence, there exist two arrows $p_1\colon \tproi{D}{1}\to \tpro{D}$, $p_2\colon \tproi{D}{2}\to \tpro{D}$ such that, for every $X\in  \Delta(\der{D}_1)$ and $Y\in  \Delta(\der{D}_2)$
		\[p_1\circ \iota_{1, X} = \iota_X \qquad p_2\circ \iota_{2, Y}=\iota_{2,Y}\]
		In particular, this entails the commutativity of the square
				\[\xymatrix{H \ar[dr]^{\iota_H} \ar[r]^-{\iota_{2, H}} \ar[d]_-{\iota_{1, H}} & \tproi{D}{2} \ar[d]^{p_2}\\  \tproi{D}{1} \ar[r]_{p_1}& \tpro{D}}\]

@@

	For uniqueness, suppose that $c'\colon \tpro{D}\to C$ is such that
	\[c'\circ p_1=a \qquad c'\circ p_2 = b\]
Then, for every $X\in \Delta(\der{D})$ we have
\begin{align*}
	c'\circ \iota_X &= \begin{cases}
	c'\circ p_1\circ \iota_{1, X} & X\in \Delta(\der{D}_1)\\
	c'\circ p_2\circ \iota_{2, X} & X\in \Delta(\der{D}_2)
	\end{cases}\\&=\begin{cases}
a\circ \iota_{1, X} & X\in \Delta(\der{D}_1)\\
b\circ \iota_{2, X} & X\in \Delta(\der{D}_2)
	\end{cases}\\&=c_X\\&=c\circ \iota_X
\end{align*}
showing that $c'=c$ as wanted.

\item Let $a\colon \lpro \der{\hat{D}}\rpro\to C$ , $b\colon \lpro \der{D}' \rpro\to C$ be two arrows fitting in the diagram 
	\[\xymatrix{D_{j} \ar[ddrr]^{\iota_{D_j}}\ar[r]^{g_j}\ar[d]_{f_j}& G_{j+1} \ar[r]^{\hat{\iota}_{G_{j+1}}} & \lpro \der{\hat{D}}\rpro \ar@/^.4cm/[dddr]^{a} \ar[dd]^{p_2} \\ G_j \ar[d]_{\iota'_{G_j}}\\ \lpro \der{D}' \rpro \ar@/_.4cm/[rrrd]_{b} \ar[rr]_{p_1}  &&\tpro{D} \ar@{.>}[dr]^{c}\\ &&& C }\]

	We can define a cocone $(C, \{c_X\}_{X\in \Delta(\der{D})})$ putting
		\[c_X:=\begin{cases}
		a\circ \hat{\iota}_{ X} & X\in \Delta(\der{\hat{D}})\\
		b\circ \iota'_{ X} & X\in \Delta(\der{D}')\\
		b\circ \iota'_{G_j} \circ f_j & X=D_j
	\end{cases}\]
	Thus we get a unique arrow $c\colon \tpro{D}\to C$ such that 
	\[c\circ \iota_X = c_X\]
	Now, for every $X\in  \Delta(\der{D}')$ and $Y\in  \Delta(\der{\hat{D}})$ we have
	\[\begin{split}
		c\circ p_1 \circ \iota'_{X}&=c\circ \iota_{X}\\&=c_X \\&=b\circ \iota'_{X}\\&=b\circ p_1\circ \iota'_{X}
	\end{split}\qquad \begin{split}
		c\circ p_2 \circ \hat{\iota}_{Y}&=c\circ \iota_{Y}\\&=c_Y \\&=a\circ \hat{\iota}_{Y}\\&=a\circ p_2\circ \hat{\iota}_{Y}
	\end{split}\]
	so that we can deduce
	\[c\circ p_1=a \qquad c\circ p_2 = b\]

>>>>

\newcommand{\pgfBox}{
  \begin{pgfonlayer}{background} 
    \fill[blue!2,thick,draw=black!50,rounded corners,inner sep=3mm] ([xshift=-1.5pt,yshift=-1.5pt]current bounding box.south west) rectangle ([xshift=1.5pt,yshift=1.5pt]current bounding box.north east);
  \end{pgfonlayer}
}
\usepackage{scalerel}
\newcommand{\smallmin}{\scaleobj{0.6}{-}}
\newcommand{\Deltamin}{\Delta^{\hspace{-1pt}\downarrow\hspace{-1pt}}}
\newcommand{\Rrel}[1]   {\stackrel{{#1}}{\Longrightarrow}}
\newcommand{\oa}{\overline a}
\newcommand{\ob}{\overline b}
\newcommand{\oc}{\overline c}
\newcommand{\od}{\overline d}
\newcommand{\rec}{\emph{rec}}
\newcommand{\fn}[1]{{\mathtt{fn}}(#1)}
%\usepackage{latexsym}
\usepackage{stmaryrd}
\def\encodep#1{\llfloor#1\rrfloor}

@@

\begin{definition}
	Let $(\X, \R)$ be a DPO-rewrityng system, with $\X$ $\mathcal{M}$-adhesive. The \emph{DPO-derivation graph} of $(\X, \R)$ is the (large)  directed graph $\gpo$ having as vertices the objects of $\X$ and in which an edge between $G$ and $H$ is a direct derivation $\dder{D}\colon G\Mapsto H$.	A \emph{derivation} $\der{D}$ between two objects $G$ and $H$ is a path between them in $\gpo$. The \emph{source} and \emph{target} of $\der{D}$ are, respectively, $G$ and $H$.
\end{definition}

@@

\begin{notation}
	We will extend the use of the words length, source and target to decorated derivations in the obvious way, forgetting the decorations $\alpha$ and $\omega$.
\end{notation}

@@

\begin{definition}Let $(\X, \R)$ be a DPO-rewriting system,  an \emph{abstraction equivalence} between two derivations $\der{D}$ and $\der{D'}$ with the same length and such that $r(\der{D})=r(\der{D}')$, is a family of isomorphisms $\{\phi_X\}_{X\in \Deltamin(\der{D})}$ such that, for every $i\in [0, \lgh(\der{D})]$ the following diagram commutes
	\[\xymatrix@C=40pt{G'_i&D'_i \ar[r]^{g'_i} \ar[l]_{f'_i}&G'_{i+1}\\  L_i \ar[u]^{n'_i} \ar[d]_{n_i}& K_i \ar[u]^{k'_i} \ar[d]_{k_i} \ar[r]^{r_i} \ar[l]_{l_i} &R\ar[u]^{h'_i} \ar[d]_{h_i}\\G_i \ar@/_.45cm/[uu]_(.35){\phi_{G_i}}|\hole&D_i\ar[l]^{f_i}\ar@/_.45cm/[uu]_(.35){\phi_{D_i}}|\hole \ar[r]_{g_i}&G_{i+1}\ar@/_.45cm/[uu]_{\phi_{H_i}}}\]

@@

Given DPO-rewriting system $(\X, \R)$,  we have already noted in \Cref{rem:func} that a derivation $\der{D}$  determines a diagram $\Delta(\der{D})$ in $\X$. We can then wonder if such a diagram has a colimit. Clearly if $\der{D}$ is the empty derivation $G$ then a colimit for $\Delta(\der{D})$ is simply the object $G$. More generally, we have the following result.

\begin{lemma}\label{lem:colim}
	Let $\X$ be an $\mathcal{M}$-adhesive category and $(\X, \R)$ a left-linear DPO-rewriting system over it. The following properties hold true.
	\begin{enumerate}
		\item  If $\der{D}$ is a derivation from $G$ to $H$, then the diagram $\Delta(\der{D})$ has a colimit $(\tpro{D}, \{\iota_X\}_{X\in \Delta(\der{D})})$ such that $\iota_H$ belongs to $\mathcal{M}$.
		\item Let $\der{D}$ be the concatenation $\der{D}_1\cdot \der{D}_2$ of two derivations $\der{D}_1=\{\dder{D}_{1,i}\}_{i=0}^{n_1}$ between $G$ and $H$ and $\der{D}_2=\{\dder{D}_{2,j}\}_{j=0}^{n_2}$ between $H$ and $T$,  then the colimiting cocone $(\tpro{D}, \{\iota_X\}_{X\in \Delta(\der{D})})$ exists too and there is a pushout square
		\[\xymatrix{H\ar[r]^-{\iota_{2, H}} \ar@{>->}[d]_-{\iota_{1, H}} & \tproi{D}{2} \ar[d]^{p_2}\\  \tproi{D}{1} \ar[r]_{p_1}& \tpro{D}}\]
		where $(\tproi{D}{1}, \{\iota_{1, X}\}_{X\in \Delta(\der{D}_1)})$ and $(\tproi{D}{2}, \{\iota_{2, X}\}_{X\in \Delta(\der{D}_2)})$ are the colimiting cocone for $\Delta(\der{D}_1)$ and $\Delta(\der{D}_2)$, respectively.
	\end{enumerate}
\end{lemma}
\begin{remark}\label{rem:cof}
Let $I:\Deltamin(\dder{D})\to \Delta(\dder{D})$ be the inclusion functor. It is immediate to see that such functor is \emph{final} \cite{mac2013categories}. This means that for every functor $F\colon \Delta(\dder{D})\to \Y$ we have:
\begin{enumerate}
	\item if  $(C, \{c_X\}_{X\in \Deltamin(\dder{D})})$ is colimiting for $F\circ I$, then there exists a colimiting cocone $(D, \{d_X\}_{X\in \Delta(\dder{D})})$ for $F$;
	\item $(C, \{c_X\}_{X\in \Deltamin(\dder{D})})$ and $(D, \{d_X\}_{X\in \Delta(\dder{D})})$ are colimiting for, repsectively, $F\circ I$ and $F$, then the canonical arrow $\phi\colon C\to D$ induced by $(D, \{d_X\}_{X\in \Deltamin(\dder{D})})$ is an isomorphism.
\end{enumerate}
\end{remark}

\begin{proof}\begin{enumerate}
		\item Let us proceed by induction on the length of $\der{D}$.
		\begin{itemize}
			\item  $\lgh(\dder{D})=0$. then the $\tpro{\dder{D}}$ is simply $(G, \{\id{G}\})$ and $\id{G}\in \mathcal{M}$.¸
			\item $\lgh(\dder{D})=0$. Suppose that $\dder{D}$ has as its single component the derivation
			\[\xymatrix{L \ar[d]_{m}& K \ar[d]^{k}\ar@{>->}[l]_{l} \ar[r]^{r} & R \ar[d]^{h} \\G& \ar@{>->}[l]^{f} D \ar[r]_{g}& H  \\}\]
			The arrow $f$ is  the pushout of $l$ and so it is in in $\mathcal{M}$. We can thus consider the $\mathcal{M}$-pushout square
			\[\xymatrix{D \ar@{>->}[d]_{f} \ar[r]^{g} & H \ar@{>->}[d]^{p} \\G \ar[r]_{q}& P }\]
			Since $p\in \mathcal{M}$, the thesis follows immediately from \Cref{rem:cof}. 
			\item $\lgh(\dder{D})\geq 2$. Let $\der{D}$ be $\{\dder{D}_i\}_{i=0}^n$ with $n\geq 1$. Let also $\der{D}'$ be $\{\dder{D}_i\}^{n-1}_{i=0}$ and $\rho_n=(l_n, r_n)$ be the rule applied in $\dder{D}_n$. The pushout of $l_n$ is the arrow $f_n\colon D_n\to G_n$ is in $\mathcal{M}$ and, by inductive hypothesis, $\iota_{G_{n}}\colon G_{n}\to \lpro \der{D}'\rpro$ is in $\mathcal{M}$ too. Thus, we can consider the diagram below, having a pushout as its lower half.
			\[\xymatrix{L_n \ar[d]_{m_{n}}& K_{n} \ar[d]^{k_{n}}\ar@{>->}[l]_{l_{n}} \ar[r]^{r_{n}} & R_{n} \ar[d]^{h_n} \\G_{n} \ar@{>->}[d]_{\iota'_{G_{n}}}& \ar@{>->}[l]^{f_n} D_n \ar[r]_{g_n}& H  \ar@{>->}[d]^{q}\\ \lpro \der{D}' \ar[rr]_{p}\rpro && P}\] 
			Notice that, as in the point above, the arrow $q\colon H\to P$ is the pushout of an element in $\mathcal{M}$, therefore it is enough to show that the diagram so constructed provides a colimiting cocone for $\Delta(\der{D})$.

			Let $(C, \{c_X\}_{X\in \Delta(\der{D})})$ be a cocone, since $\Delta(\der{D}')$ is a subdiagram of $\Delta(\der{D})$, we get another cocone $(c, \{c_X\}_{X\in \Delta(\der{D}')})$ which induces an arrow $c'\colon \lpro \der{D}' \rpro \to C$ such that
			\begin{align*}
				c'\circ \iota_{G_n} \circ f_n &=c_{G_n} \circ f_n\\&= c_{D_n}\\&= c_{H}\circ g_n
			\end{align*}
			Therefore the arrows $c'$ and $c_H$ induce a morphism $c\colon P\to C$ and the thesis now follows at once.
		\end{itemize}

		\item  As a first step, notice that $(\tpro{D}, \{\iota_X\}_{X\in \Delta(\der{D}_1)})$ and $(\tpro{D}, \{\iota_X\}_{X\in \Delta(\der{D}_2)})$ are cocone on, respectively, $\Delta(\der{D}_1)$ and $\Delta(\der{D}_2)$. Hence, there exist two arrows $p_1\colon \tproi{D}{1}\to \tpro{D}$, $p_2\colon \tproi{D}{2}\to \tpro{D}$ such that, for every $X\in  \Delta(\der{D}_1)$ and $Y\in  \Delta(\der{D}_2)$
		\[p_1\circ \iota_{1, X} = \iota_X \qquad p_2\circ \iota_{2, Y}=\iota_{2,Y}\]
		In particular, this entails the commutativity of the square
				\[\xymatrix{H \ar[dr]^{\iota_H} \ar[r]^-{\iota_{2, H}} \ar@{>->}[d]_-{\iota_{1, H}} & \tproi{D}{2} \ar[d]^{p_2}\\  \tproi{D}{1} \ar[r]_{p_1}& \tpro{D}}\]

@@

	For uniqueness, suppose that $c'\colon \tpro{D}\to C$ is such that
	\[c'\circ p_1=a \qquad c'\circ p_2 = b\]
Then, for every $X\in \Delta(\der{D})$ we have
\begin{align*}
	c'\circ \iota_X &= \begin{cases}
	c'\circ p_1\circ \iota_{1, X} & X\in \Delta(\der{D}_1)\\
	c'\circ p_2\circ \iota_{2, X} & X\in \Delta(\der{D}_2)
	\end{cases}\\&=\begin{cases}
a\circ \iota_{1, X} & X\in \Delta(\der{D}_1)\\
b\circ \iota_{2, X} & X\in \Delta(\der{D}_2)
	\end{cases}\\&=c_X\\&=c\circ \iota_X
\end{align*}
showing that $c'=c$ as wanted.	 \qedhere 
	\end{enumerate}
\end{proof}

\begin{corollary}\label{cor:colim}
Let $\der{D}=\{\dder{D}_i\}_{i=0}^n$ a derivation of length $n+1$ and fix an index $j\in[0,n]$. Define
\[\der{D}^j_1:=\{\dder{D}_i\}_{i=0}^{j-1} \qquad  \der{D}^j_2=\{\dder{D}_j\} \qquad \der{D}^j_3:=\{\dder{D}_i\}_{i=j+1}^n\]
with the convention that $\der{D}^0_1$ and $\der{D}^n_3$ are the empty derivation on, respectively, $G_0$ and $G_n$. Then the square below is a pushout and a pullback
\[\xymatrix@C=30pt{D_{j} \ar[ddrr]^{\iota_{D_j}} \ar[r]^{g_j}\ar@{>->}[d]_{f_j}& G_{j+1} \ar[ddr]^{\iota_{G_{j+1}}} \ar[r]^-{\iota_{3, G_{j+1}}} & \lpro \der{D}^j_3\rpro \ar[dd]^{p_2} \\ G_j\ar[drr]_{\iota_{G_{j}}} \ar@{>->}[d]_{\iota_{1,G_j}}\\ \lpro \der{D}^j_1 \rpro \ar[rr]_{p_1}  &&\tpro{D} }\] 
Where the two arrows $p_1\colon \lpro\der{D}^j_1 \rpro\to \tpro{D}$, $p_2\colon \lpro \der{D}^j_3\rpro \to \tpro{D}$ are induced by the cocones $(\tpro{D}, \{\iota_{X}\}_{X\in \Delta(\der{D}^j_1)})$ and $(\tpro{D}, \{\iota_{X}\}_{X\in \Delta(\der{D}^j_3)})$, respectively.
\end{corollary}
\begin{remark}
If $\der{D}$ is empty then $\der{D}^j_1, \der{D}^j_2$ and $\der{D}^j_3$ are empty too.
\end{remark}
\begin{proof}
We can notice that $\der{D}=\der{D}^j_1\cdot \der{D}^j_2 \cdot \der{D}^j_3$. By the first and the second point of \Cref{lem:colim} then we get the following diagram, in which all squares are $\mathcal{M}$-pushouts.

\[\xymatrix@C=30pt{D_{j}  \ar[r]^{g_j}\ar@{>->}[d]_{f_j}& G_{j+1}  \ar[r]^-{\iota_{3,G_{j+1}}} \ar[d]_{\iota_{2, G_{j+1}}}  \ar@/^.5cm/[dd]^{\iota_{1,2, G_{j+1}}}& \lpro \der{D}^j_3\rpro \ar[dd]^{p_2} \\ G_j \ar[r]^{\iota_{2, G_j}}\ar@{>->}[d]_{\iota_{1,G_j}} & \lpro \der{D}^j_2\rpro \ar[d]_a\\ \lpro \der{D}^j_1 \rpro \ar@/_.4cm/[rr]_{p_1} \ar[r]^b &\lpro \der{D}^j_1\cdot \der{D}^j_2 \rpro \ar[r]^c&\tpro{D} }\] 
Applying \Cref{lem:po1} twice we get that the whole square is an $\mathcal{M}$-pushout. Then the thesis follows from \Cref{prop:pbpoad}.
\end{proof}

\begin{remark}\label{rem:zero1} In particular, considering $j=0$ or $j=n$, we get that the following two squares are $\mathcal{M}$-pushouts and, thus, pullbacks.
	\[\xymatrix{D_0 \ar[r]^-{\iota_{3, D_0}} \ar@{>->}[d]_{f_0}& \lpro \der{D}^0_3 \rpro \ar@{>->}[d]^{p_2} &D_n \ar@{>->}[d]_{\iota_{1, D_n}}\ar[r]^{g_n}& G_n \ar@{>->}[d]^{\iota_{G_n}}\\ G \ar[r]_{\iota_G}& \tpro{D} & \lpro \der{D}^n_1 \rpro \ar[r]_{p_1}  & \tpro{D}}\]
\end{remark}

\begin{corollary}\label{cor:ele}
	Let $\der{D}=\{\dder{D}_{i}\}_{i=0}^n$ be a derivation between $G$ and $H$. Let $j$ and $k$ be two indexes less or equal than $n+1$ and suppose that $j< k$.  Consider two arrows $a\colon T\to G_j$, $b\colon T\to G_k$. If $\iota_{G_j}\circ a = \iota_{G_k}\circ b$, 
	then  there exist a unique arrow $c\colon T\to D_j $  such that \[f_j\circ c = a\qquad \iota_{D_j}\circ c =\iota_{G_k}\circ b\]
	\end{corollary}
\begin{proof} Consider the diagram
			\[\xymatrix@C=34pt{T  \ar@/_.6cm/[dd]_{a}\ar@{.>}[d]^{c}\ar[rr]^{b}&& G_k \ar[d]_{\iota_{3, G_k}} \ar@/^.6cm/[ddd]^{\iota_{G_k}}\\D_{j} \ar[r]^{g_j} \ar[ddrr]^{\iota_{D_j}}\ar@{>->}[d]^{f_j}& G_{j+1} \ar[r]^-{\iota_{3,G_{j+1}}} \ar[ddr]^{\iota_{G_{j+1}}} & \lpro \der{D}^j_3\rpro \ar[dd]_(.6){p_2} \\ G_j \ar@{>->}[d]_{\iota_{1,G_j}} \ar[drr]_{\iota_{G_j}}\\ \lpro \der{D}^j_1 \rpro \ar[rr]_{p_1}  &&\tpro{D} }\] 
			Thanks to \Cref{cor:colim} we know that the bottom right rectangle in the diagram above is a pullback and the thesis follows at once.
\end{proof}
